---
title: "Organizational Graphs"
output:
  html_document:
    df_print: paged
---

### Visualizing Organizations using iGraph   

#### Read in the data  
```{r}
library(igraph)
library(network)
nodes <- read.csv("~/rdata/nodes.csv", header=T)
links <- read.csv("~/rdata/edges.csv", header=T, as.is=T)
```

#### Verify we have what we expect    

1. Number of nodes & edges  
  
```{r}
nrow(nodes); length(unique(nodes$id))
nrow(links); nrow(unique(links[,c("from", "to")]))
```
2. Glimpse into nodes   
```{r}
head(nodes); summary(nodes)
```
3. Glimpse into edges       
```{r}
head(links)
```
#### Clean up and prepare the graph data         
   
1. Trim any whitespace   

```{r}
links$from <- trimws(links$from); links$to <- trimws(links$to); 
links$lvl <- trimws(links$lvl)
nodes$id <- trimws(nodes$id); nodes$role <- trimws(nodes$role);
```

2.  Change column 3 name to "weight" in *links* data frame   
  
```{r}
colnames(links)[3] <- "weight"
head(links)
```
3. Remove Row Names in *links* data frame      
  
```{r}
rownames(links) <- NULL  
```

4. Create a graph from the data frame     
  
```{r}
net <- graph.data.frame(links, vertices=nodes, directed=T)
```

5. Verify the edges and vertices of the graph   

```{r}
E(net);       # The edges of the "net" object
V(net);      # The vertices of the "net" object
```

6. Simplify the network matrix   

```{r}
net <- simplify(net, remove.multiple = F, remove.loops = T) 
```

7. Set the Vertex Options   

```{r}
V(net)$size <- V(net)$weight*0.6
V(net)$role <- as.factor(V(net)$role)
V(net)$label <- V(net)$name
V(net)$label.cex <- .5
V(net)$shape <- "circle"
# One of “none”, "circle", “square”, “csquare”, “rectangle”, “crectangle”, “vrectangle”, “pie”, “raster”, or “sphere”
V(net)$label.family <- "Helvetica"  # Font family of the label (e.g.“Times”, “Helvetica”)
V(net)$size <- 20  
V(net)$font <- 2 # Font: 1 plain, 2 bold, 3, italic, 4 bold italic, 5 symbol
V(net)$distance <- 20 #Distance between the label and the vertex
V(net)$degree <- 0 #0 right, “pi” is left, “pi/2” is below, and “-pi/2” is above

```

8. Set the Edge Options   
```{r}
# Set edge width based on weight:
E(net)$width <- as.numeric(E(net)$weight)/6
#change arrow size and edge color:
E(net)$arrow.size <- .4
E(net)$edge.color <- "gray80"
E(net)$width <- 1+as.numeric(E(net)$weight)/12
E(net)$curved <- .5
```

9. Quick Plot   

```{r}
plot(net, layout=layout.davidson.harel(net))
```

#### Add Attributes      

1. Add Vertex Attributes  
```{r}
# Assign ID to "name" attribute
V(net)$name <- nodes$id
# Assign role to "role" attribute
V(net)$role <- nodes$role
vertex_attr(net)
```

2. Add Edge Attributes   
  
```{r}
# Assign role hierarchical level to "type" attribute  
E(net)$type <- ifelse(links$weight==1, "Individual Contributer", 
                      ifelse(links$weight==2, "Manager",
                             ifelse(links$weight==3, "Director", "VP")))
edge_attr(net)
```

3. Add Graph Attributes   
  
```{r}
# Assign "Reporting Network" to the "name" attribute of the graph
net <- set_graph_attr(net, "name", "Reporting Network")

# Assign "Product" to the "org" attribute of the graph
net <- set_graph_attr(net, "org", "Product")

# To demonstrate deleting attributes:  
# net <- set_graph_attr(net, "something", "A thing")
# net <- delete_graph_attr(net, "something")

graph_attr_names(net)
graph_attr(net)
```
  
#### Plot Graph       
  
```{r fig.width=12, fig.height=7}
V(net)$colors <- nodes$lvl
library(RColorBrewer)
darkcols <- brewer.pal(4, "Dark2")

l <- layout_with_dh(net)

plot(net, 
     layout=l*1,
     edge.arrow.size=.3,
     margin = 0,
     vertex.color=darkcols[V(net)$colors],
     vertex.size=20,
     vertex.label=V(net)$role,
     vertex.frame.color="gray", 
     vertex.label.color="black", 
     vertex.label.cex=0.8, 
     vertex.label.dist=0, 
     edge.curved=0.2) 
legend(x=-1.5, y=-1.1, c("Individual Contributer","Manager","Director", "VP"),
       pch=21, col="#777777", pt.bg=darkcols, pt.cex=1, cex=.8, bty="n", ncol=1)

```

12. Gather some details about the structure of the graph  

```{r}
edge_density(net, loops=F)
ecount(net)/(vcount(net)*(vcount(net)-1))  #for a directed network
diameter(net, directed=F, weights=NA)
diameter(net, directed=F)

diam <- get_diameter(net, directed=T)
diam
```

13. Plot where vertex size indicates network size   

```{r}
deg <- degree(net)
plot(net, vertex.size=deg*3)
```



